---
title: |
  | IRT model: AS index for 2013
  | Sustainability items with 4 categories 
  | Exploring AS scores estimated with incomplete items
  | 2 out of 9 items missing for 30 percent of the units
author: "Christoph Wunder"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab
    highlight: pygments
    toc: yes
    toc_depth: 4
    number_sections: yes
    toc_float:
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '2'
# bibliography: /Users/Christoph/Documents/Uni/Literatur/lit_DB/litdb_sp.bib
# biblio-style: apalike
# link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, warning=FALSE)
rm(list=ls())
library(tidyverse)
library(haven)
library(brms)
library(bayesplot)
library(ggplot2)
theme_set(theme_bw())
library(tictoc)
library(sampling)
library(shinystan)
```

We explore the estimation of the AS index using incomplete data on items. For that purpose, we compare AS scores that are estimated with incomplete data on items with the AS scores that are estimated on the complete data. The incomplete data set considers one to three missing items on 10%, 30%, and 50% of the units.  

# Data

The data are prepared in `/stata/cmd-cs/11-data_processing-cs.do`.

* 2013 cross section data
* We fit a model using all observations with all (9) items. This is the control model.

```{r}
d <- read_dta("../stata/dat/200-asi-ord4-2013master.dta")
d <- rename(d, id = idn)
d$id <- as.integer(d$id)
d$item <- as.factor(d$item)
# d <- d %>% filter( item == c("1", "2", "3") )
table(d$item)
```

```{r}
d$y <- factor( d$y+1, 
               levels = c("1", "2", "3", "4"), 
               labels = c("very unsustainable", "unsustainable", "sustainable", "very sustainable"),
               ordered = TRUE ) 
glimpse(d, width = 60)
table(d$y)
```


Select subsample of data set:

```{r, eval=FALSE}
set.seed(647382)
n <- round(length(unique(d$id))*0.05, 0)
samplecluster <- cluster(d, clustername=c("id"), size=n, method="srswor")
d <- getdata(d, samplecluster)
```

Select only required variables

```{r}
d <- d %>% select(id, item ,y)
```

We randomly drop 2 out of 9 items in 30% of the farms.

```{r}
# select 30% of the farms
set.seed(134679)
n <- round(length(unique(d$id))*0.3, 0)
samplecluster <- cluster(d, clustername=c("id"), size=n, method="srswor")
units2drop <- getdata(d, samplecluster)
units2drop <- units2drop %>% select(id, item)
units2drop <- units2drop[order(units2drop$id, units2drop$item),]

# select 2 different items out of 9 items
length(unique(units2drop$id))
items2drop <- data.frame( id = unique(units2drop$id),
                          dropitem1 = sample(1:9, n, replace = TRUE))

for (i in 1:nrow(items2drop)) {
  items2drop$dropitem2[i] = sample(setdiff(1:9, items2drop$dropitem1[i]), 1, replace = TRUE)
}

# check if the same item was dropped
sum(items2drop$dropitem1 == items2drop$dropitem2)

# merge items2drop and units2drop
d2drop <- left_join(units2drop, items2drop, by = "id")

# merge d2drop with original data frame d
d <- left_join(d, d2drop, by = c("id", "item"))
rm(d2drop, n, samplecluster)

# make data incomplete
d$y <- ifelse( is.na(d$dropitem1), d$y,
               ifelse( d$dropitem1 == d$item, NA,
                       ifelse( d$dropitem2 == d$item, NA, d$y )))

d %>% filter(!is.na(dropitem1)) %>% arrange(id, item) %>% head(27)
```


```{r}
summary(d)
```

# Fitting the model in brms

```{r, message=TRUE, warning=TRUE}
tic("fit1")
prior <-
  # discrimination
  prior("normal(0, 2)", class = "Intercept", dpar = "disc") + # +             # alpha
  prior("normal(0, 2)", class = "sd", group = "item", dpar = "disc") +    # sigma_alpha
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 2)", class = "sd", coef = "Intercept", group = "item") 
formula <- brmsformula(
  # y | thres(gr = item) ~ 1 + (1|id),
  y ~ 1 +  (1|i|item) + (1|id),
  disc ~ 1 + (1|i|item)
)
fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 4,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "../results/98-cs-2013-incomplete-2-30.rds"
)
toc()
```

```{r}
summary(fit1)
```

Extract diagnostic quantities:

```{r}
# help("diagnostic-quantities")

# extract the number of divergence transitions
np <- nuts_params(fit1)
sum(subset(np, Parameter == "divergent__")$Value)
```

```{r}
# launch_shinystan(fit1)
```


```{r}
prior_summary(fit1)
```

# Extract sustainability index (farm parameters)

```{r}
df.re <- ranef(fit1) 
dim(df.re$id) # dimension: no of units X no sum. stats (4) X no. od parameters
df.re <- data.frame( asi = df.re$id[,1,] ) # extract mean
dim(df.re)
# convert row names into first column
df.re <- tibble::rownames_to_column(df.re, "id")
head(df.re)
```

Save sustainability indicator:

```{r}
nrow(df.re) # number of AS values: one value for each farm in each year 
write.csv(df.re, file = "../results/98-asi-missing-2-30.csv")
```

# Compare with AS score estimated from complete data

```{r}
d.complete <- read.csv("../results/90-asi.csv", header = TRUE)
d.complete <- d.complete %>% select( -X )
d.complete <- d.complete[order(d.complete$id),]
head(d.complete)

d.incomplete <- read.csv("../results/98-asi-missing-2-30.csv")
d.incomplete <- d.incomplete %>% select( -X )
d.incomplete <- d.incomplete[order(d.complete$id),]
head(d.incomplete)

d.combine <- left_join(d.complete, d.incomplete, by = "id")
```

```{r}
summary(lm(asi.x ~ asi.y, data = d.combine))
```
```{r}
cor(d.combine$asi.x, d.combine$asi.y)
```

# Original computing environment {-}

```{r}
devtools::session_info()
```

```{r}
knitr::knit_exit()
```
