---
title: |
  | IRT model: AS index for 2013
  | Sustainability indicators with 4 categories 
author: "Christoph Wunder"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab
    highlight: pygments
    toc: yes
    toc_depth: 4
    number_sections: yes
    toc_float:
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '2'
# bibliography: /Users/Christoph/Documents/Uni/Literatur/lit_DB/litdb_sp.bib
# biblio-style: apalike
# link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, warning=FALSE)
rm(list=ls())
library(tidyverse)
library(haven)
library(brms)
library(bayesplot)
library(ggplot2)
theme_set(theme_bw())
library(tictoc)
library(sampling)
library(shinystan)
```


# Data

The data are prepared in `/stata/cmd-cs/11-data_processing-cs.do`.

* 2013 cross section data
* We fit a model using all observations with all (9) items. This is the control model.

```{r}
d <- read_dta("../stata/dat/200-asi-ord4-2013master.dta")
d <- rename(d, id = idn)
d$id <- as.integer(d$id)
d$item <- as.factor(d$item)
# d <- d %>% filter( item == c("1", "2", "3") )
table(d$item)
```


```{r}
d$y <- factor( d$y, levels = c("0", "1", "2", "3"), ordered = TRUE ) 
glimpse(d, width = 60)
table(d$y)
```

Select subsample of data set:

```{r, eval=FALSE}
set.seed(647382)
n <- round(length(unique(d$id))*0.05, 0)
samplecluster <- cluster(d, clustername=c("id"), size=n, method="srswor")
dsample <- getdata(d, samplecluster)
d <- dsample
```

```{r}
summary(d)
```

```{r}
# cbind(n, nrow(d), nrow(d)/n)
```

```{r}
ggplot( d, aes(x=y) ) +
  stat_count(width = 0.5) +
  xlab("") +
  facet_wrap(~item)
```

# Fitting the model in brms

```{r, message=TRUE, warning=TRUE}
tic("fit1")
prior <-
  # discrimination
  prior("normal(0, 2)", class = "Intercept", dpar = "disc") + # +             # alpha
  prior("normal(0, 2)", class = "sd", group = "item", dpar = "disc") +    # sigma_alpha
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 2)", class = "sd", coef = "Intercept", group = "item") 
formula <- brmsformula(
  # y | thres(gr = item) ~ 1 + (1|id),
  y ~ 1 +  (1|i|item) + (1|id),
  disc ~ 1 + (1|i|item)
)
fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 4,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "../results/90-1-cross-sec-2013.rds"
)
toc()
```

```{r}
summary(fit1)
```

Extract diagnostic quantities:

```{r}
# help("diagnostic-quantities")

# extract the number of divergence transitions
np <- nuts_params(fit1)
sum(subset(np, Parameter == "divergent__")$Value)
```

```{r}
# launch_shinystan(fit1)
```


```{r}
prior_summary(fit1)
```

# Interpretation

## Easiness and discrimination parameters

```{r}
# extract item and easiness parameters
ranef1 <- ranef(fit1)
# discriminations
alpha <- ranef1$item[, , "disc_Intercept"] %>%
  exp() %>% 
	as_tibble() %>%
	rownames_to_column()
# item easinesses (deviations from thresholds)
beta <- ranef1$item[, , "Intercept"] %>%
	as_tibble() %>%
	rownames_to_column()
# put easinesses and discriminations together
bind_rows(beta, alpha, .id = "nlpar") %>%
	rename(item = "rowname") %>%
	mutate(item = as.numeric(item)) %>%
	mutate(nlpar = factor(nlpar, labels = c("Easiness", "Discrimination"))) %>%
	ggplot(aes(item, Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap("nlpar", scales = "free_x") +
	geom_pointrange() +
	coord_flip() +
	labs(x = "Item Number") + 
  scale_x_discrete( limits=1:9 )
```

# Extract sustainability index (farm parameters)

## Random unit-specific and time-varying effects

```{r}
df.re <- ranef(fit1) 
dim(df.re$id) # dimension: no of units X no sum. stats (4) X no. od parameters
df.re <- data.frame( asi = df.re$id[,1,] ) # extract mean
dim(df.re)
# convert row names into first column
df.re <- tibble::rownames_to_column(df.re, "id")
head(df.re)
```


Save sustainability indicator:

```{r}
nrow(df.re) # number of AS values: one value for each farm in each year 
write.csv(df.re, file = "../results/90-1-asi.csv")
```


# Original computing environment {-}

```{r}
devtools::session_info()
```


```{r}
knitr::knit_exit()
```
