---
title: |
  | IRT model: AS index for 2013
  | Sustainability items with 4 categories 
  | Scale linking: East and West Germany
author: "Christoph Wunder"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab
    highlight: pygments
    toc: yes
    toc_depth: 4
    number_sections: yes
    toc_float:
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '2'
# bibliography: /Users/Christoph/Documents/Uni/Literatur/lit_DB/litdb_sp.bib
# biblio-style: apalike
# link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(haven)
library(brms)
library(bayesplot)
library(ggplot2)
theme_set(theme_bw())
library(tictoc)
library(sampling)
#library(shinystan)
```

To test the argument for scale linking, we divide the data set into two parts (East Germany and West Germany) and delete one item in each part. 


# Data

The data are prepared in `stata/cmd-cs/11-data_processing-cs.do`.

```{r data}
d <- read_dta("../../stata/dat/200-asi-ord4-2013grouped.dta")
d <- rename(d, id = idn)
d$id <- as.integer(d$id)
d$item <- as.factor(d$item)
table(d$item)
```
Select subsample of data set:

```{r subsample, eval=FALSE}
set.seed(647382)
n <- round(length(unique(d$id))*0.05, 0)
samplecluster <- cluster(d, clustername=c("id"), size=n, method="srswor")
d <- getdata(d, samplecluster)
```

Select only required variables

```{r select variables}
d <- d %>% select(id, item , y, group, NUTS2)
```

We split the data set into two parts.

```{r east-west divide}
# group == 1 is West Germany, group == 2 is East Germany
d$region <- ifelse( d$group==1, 1, 0 )
d$region <- factor( d$region, levels = c(0, 1), labels = c("East", "West") )
table(d$region, useNA = "ifany")
```
```{r create missing items}
d$y <- ifelse( d$region == "West" & d$item == west.item, NA,
                   ifelse( d$region == "East" & d$item == east.item, NA, d$y ))
```

```{r recode items}
d$y <- factor( d$y+1, 
               levels = c("1", "2", "3", "4"), 
               labels = c("very unsustainable", "unsustainable", "sustainable", "very sustainable"),
               ordered = TRUE ) 
glimpse(d, width = 60)
table(d$y)
```

```{r}
xtabs( ~ item + y, data=d, subset=region=="East")
```

```{r}
xtabs( ~ item + y, data=d, subset=region=="West")
```


# Fitting the model in brms

```{r fit, message=TRUE, warning=TRUE}
tic("fit1")
prior <-
  # discrimination
  prior("normal(0, 2)", class = "Intercept", dpar = "disc") + # +             # alpha
  prior("normal(0, 2)", class = "sd", group = "item", dpar = "disc") +    # sigma_alpha
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 2)", class = "sd", coef = "Intercept", group = "item") +
  # covariate effects 
  prior("normal(0, 2)", class = "b") 
formula <- brmsformula(
  y ~ 1 + NUTS2 + (1|i|item) + (1|id),
  disc ~ 1 + (1|i|item)
)
fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 4,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = paste0("../../results/30-scale-linking/32-NUTS2/east-", 
                east.item, "-west-", west.item, ".rds")
)
toc()
```

```{r estimation results}
summary(fit1)
```

Extract diagnostic quantities:

```{r diagnostics}
# extract the number of divergence transitions
np <- nuts_params(fit1)
sum(subset(np, Parameter == "divergent__")$Value)
```

```{r}
# launch_shinystan(fit1)
```


```{r priors}
prior_summary(fit1)
```

# Interpretation

```{r predicted probabilities}
newx <- data.frame( NUTS2 = unique(d$NUTS2) )
pp <- fitted(fit1, newdata = newx, re_formula = NA, probs = c(0.10, 0.90))
pp4 <- pp[,,"P(Y = very sustainable)"]
cbind(newx, pp4)
```

# Original computing environment {-}

```{r}
devtools::session_info()
```

