---
title: |
  | IRT model with covariates (splines)
author: "Christoph Wunder"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab
    highlight: pygments
    toc: yes
    toc_depth: 4
    number_sections: yes
    # toc_float:
    #   smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '2'
# bibliography: /Users/Christoph/Documents/Uni/Literatur/lit_DB/litdb_sp.bib
# biblio-style: apalike
# link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, warning=FALSE)
rm(list=ls())
library(tidyverse)
library(haven)
library(brms)
library(bayesplot)
library(ggplot2)
theme_set(theme_bw())
library(tictoc)
library(sampling)
library(shinystan)
```

This file estimates the relationship between AS and energy and industrial using a non-parametric approach. A fixed effect of LFA is added.
  
# Data

The data are prepared starting in line 155 of the dofile `200-nfc-v3.do`.

```{r}
d <- read_dta("../data/master.dta")
d <- rename(d, id = idn)
d$id <- as.integer(d$id)
d$item <- as.factor(d$item)
d$lfa <- as.factor(d$lfa)
d$y <- factor( d$y, levels = c("0", "1", "2", "3"), ordered = TRUE ) 
```

Select subsample of data set:

```{r, eval=FALSE, include=FALSE}
# set.seed(647382)
# n <- round(length(unique(d$id))*0.05, 0)
# samplecluster <- cluster(d, clustername=c("id"), size = n, method="srswor")
# d <- getdata(d, samplecluster)
# glimpse(d, width = 60)
```

# Descriptive statistics 

```{r}
nrow(d)
nrow(d)/9
```

```{r}
summary(d)
```

```{r}
ggplot( d, aes(x=y) ) +
  stat_count(width = 0.5) +
  xlab("") +
  facet_wrap(~item)
```

# Fitting the 2PL ordinal model with unit-specific covariates (splines)

The model with item-specific thresholds produces unreasonable estimates for the discrimination parameter. We decide to use the restricitve assumption of equally shaped and shifted thresholds to proceed with a discrimination parameter that seems to be essential for the model. 

```{r, warning=FALSE, message=FALSE}
tic("fit1")
prior <-
  # discrimination
  prior("normal(0, 1)", class = "Intercept", dpar = "disc") +             # alpha
  prior("normal(0, 1)", class = "sd", group = "item", dpar = "disc") +    # sigma_alpha
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 1)", class = "sd", coef = "Intercept", group = "item") +
  # splines
  prior("normal(0, 2)", class = "sds") +
  # coefficients 
  prior("normal(0, 1)", class = "b")

formula <- brmsformula(
  y ~ 1 + (1|i|item) + (1|id) + lfa +
    s(industrial, bs="fs", k=6) + s(industrial, lfa, bs = "fs", k=6) +
    s(energy, bs="fs", k=6) + s(energy, lfa, bs = "fs", k=6),
  disc ~ 1 + (1|i|item)
)

fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 8,
  iter = 1000,
  warmup = 500,
  refresh = 100,
  file = "../results/131-2pl-spline-lfa.rds"
)
toc()
```

```{r}
prior_summary(fit1)
```

```{r}
summary(fit1)
```

Extract diagnostic quantities:

```{r}
np <- nuts_params(fit1)
sum(subset(np, Parameter == "divergent__")$Value)
```

# Interpretation

## Easiness and discrimination

```{r}
# extract item and easiness parameters
ranef1 <- ranef(fit1)
# discriminations
alpha <- ranef1$item[, , "disc_Intercept"] %>%
  exp() %>% 
    as_tibble() %>%
    rownames_to_column()
# item easinesses (deviations from thresholds)
beta <- ranef1$item[, , "Intercept"] %>%
    as_tibble() %>%
    rownames_to_column()
# put easinesses and discriminations together
bind_rows(beta, alpha, .id = "nlpar") %>%
    rename(item = "rowname") %>%
    mutate(item = as.numeric(item)) %>%
    mutate(nlpar = factor(nlpar, labels = c("Easiness", "Discrimination"))) %>%
    ggplot(aes(item, Estimate, ymin = Q2.5, ymax = Q97.5)) +
    facet_wrap("nlpar", scales = "free_x") +
    geom_pointrange() +
    coord_flip() +
    labs(x = "Item Number") + 
  scale_x_discrete( limits=1:9 )
```
  
## Effect of `industrial`

```{r}
conditional_effects(fit1,  effects = c("industrial:lfa"), prob = 0.8,
                    dpar = "mu", method = "posterior_linpred")
```

```{r}
x_pred <- tibble( industrial = rep(seq(0, 100, by=10), length(unique(d$lfa))), 
                  energy = 0,
                  lfa = rep(droplevels(sort(unique(d$lfa))), each=11))
y_pred <- fitted( fit1, 
                  newdata = x_pred,
                  scale = c("response"),
                  re_formula = NA,
                  probs = c(0.25, 0.75)) 
pred_df <- cbind.data.frame( y_pred, x_pred )

gr.1 <- 
  ggplot( pred_df ) + 
  geom_line( aes( x = industrial, y = `Estimate.P(Y = 0)`, color = "low") ) +
  geom_ribbon( aes( x = industrial, ymin = `Q25.P(Y = 0)`, ymax = `Q75.P(Y = 0)` ), alpha=0.1) +
  geom_line( aes( x = industrial, y = `Estimate.P(Y = 1)`, color = "mid") ) +
  geom_ribbon( aes( x = industrial, ymin = `Q25.P(Y = 1)`, ymax = `Q75.P(Y = 1)` ), alpha=0.2) +
  geom_line( aes( x = industrial, y = `Estimate.P(Y = 2)`, color = "high") ) +
  geom_ribbon( aes( x = industrial, ymin = `Q25.P(Y = 2)`, ymax = `Q75.P(Y = 2)` ), alpha=0.3) +
  geom_line( aes( x = industrial, y = `Estimate.P(Y = 3)`, color = "very high") ) +
  geom_ribbon( aes( x = industrial, ymin = `Q25.P(Y = 3)`, ymax = `Q75.P(Y = 3)` ), alpha=0.4) +
  facet_wrap( ~ lfa) +
  ylab("Probability") +
  scale_x_discrete("Industrial", limits = seq(0,100,10)) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position="bottom") +
  guides(color=guide_legend(title="")) 
gr.1
```

## Effect of `energy`


```{r}
conditional_effects(fit1,  effects = c("energy:lfa"), prob = 0.8,
                    dpar = "mu", method = "posterior_linpred")
```

```{r}
x_pred <- tibble( energy = rep(seq(0, 100, by=10), length(unique(d$lfa))),
                  industrial = 0,
                  lfa = rep(droplevels(sort(unique(d$lfa))), each=11))
y_pred <- fitted( fit1, 
                  newdata = x_pred,
                  scale = c("response"),
                  re_formula = NA,
                  probs = c(0.25, 0.75)) 
pred_df <- cbind.data.frame( y_pred, x_pred )

gr.2 <- 
  ggplot( pred_df ) + 
  geom_line( aes( x = energy, y = `Estimate.P(Y = 0)`, color = "low") ) +
  geom_ribbon( aes( x = energy, ymin = `Q25.P(Y = 0)`, ymax = `Q75.P(Y = 0)` ), alpha=0.1) +
  geom_line( aes( x = energy, y = `Estimate.P(Y = 1)`, color = "mid") ) +
  geom_ribbon( aes( x = energy, ymin = `Q25.P(Y = 1)`, ymax = `Q75.P(Y = 1)` ), alpha=0.2) +
  geom_line( aes( x = energy, y = `Estimate.P(Y = 2)`, color = "high") ) +
  geom_ribbon( aes( x = energy, ymin = `Q25.P(Y = 2)`, ymax = `Q75.P(Y = 2)` ), alpha=0.3) +
  geom_line( aes( x = energy, y = `Estimate.P(Y = 3)`, color = "very high") ) +
  geom_ribbon( aes( x = energy, ymin = `Q25.P(Y = 3)`, ymax = `Q75.P(Y = 3)` ), alpha=0.4) +
  facet_wrap( ~ lfa) +
  ylab("Probability") +
  scale_x_discrete("Industrial", limits = seq(0,100,10)) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position="bottom") +
  guides(color=guide_legend(title="")) 
gr.2
```

```{r}
knitr::knit_exit()
```
