---
title: |
  | IRT model without subject covariates
author: "Christoph Wunder"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab
    highlight: pygments
    toc: yes
    toc_depth: 4
    number_sections: yes
    toc_float:
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '2'
# bibliography: /Users/Christoph/Documents/Uni/Literatur/lit_DB/litdb_sp.bib
# biblio-style: apalike
# link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, warning=FALSE)
rm(list=ls())
library(tidyverse)
library(haven)
library(brms)
library(bayesplot)
library(ggplot2)
theme_set(theme_bw())
library(tictoc)
library(sampling)
library(shinystan)
```

In this file, we estimate and compare various IRT models **without** covariates. The results can be used to analyze easiness and item discrimination. 

# Data

The data are prepared starting in line 155 of the dofile `200-nfc-v3.do`.

```{r}
d <- read_dta("../data/master.dta")
d <- rename(d, id = idn)
d$id <- as.integer(d$id)
d$item <- as.factor(d$item)
d$lfa <- as.factor(d$lfa)
d$y <- factor( d$y, levels = c("0", "1", "2", "3"), ordered = TRUE ) 
```

Select subsample of data set:

```{r, eval=FALSE}
# set.seed(647382)
# n <- round(length(unique(d$id))*0.05, 0)
# samplecluster <- cluster(d, clustername=c("id"), size=n, method="srswor")
# dsample <- getdata(d, samplecluster)
# d <- dsample
```

# Descriptive statistics 

```{r}
nrow(d)
nrow(d)/9
```

```{r}
summary(d)
```

```{r}
ggplot( d, aes(x=y) ) +
  stat_count(width = 0.5) +
  xlab("") +
  facet_wrap(~item)
```

# Fitting the ordinal one-parameter IRT model 

We apply a graded response model to summarize the sustainability indicators that are measured on an ordinal four-point scale.

## Shifted thresholds

We fit the ordered one-parameter IRT model with partially pooled person and item parameters. We model an overall threshold vector that is shared across items plus an item-specific constant, implying the (restrictive) assumption that the thresholds for two items are simply shifted to the left or right relative to each other but otherwise share the same shape.

```{r, message=TRUE, warning=TRUE}
tic("fit1")
prior <-
  prior("normal(0, 3)", class = "sd", group = "id") +
  prior("normal(0, 3)", class = "sd", group = "item") 

formula <- brmsformula( y ~ 1 +  (1|item) + (1|id) )

fit1 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 8,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "../results/101-1pl.rds"
)
toc()
```

```{r}
prior_summary(fit1)
```

```{r}
summary(fit1)
```

Extract diagnostic quantities, extract the number of divergence transitions:

```{r}
np <- nuts_params(fit1)
sum(subset(np, Parameter == "divergent__")$Value)
```

## Separate threshold vectors

Next, we estimate a model where each item receives its own threshold vector. In doing so, we relax the assumption of shifted but equally shaped threshold vectors across items. Instead, we allow threshold locations and shapes to vary freely across items.

```{r, message=TRUE, warning=TRUE}
tic("fit2")
prior_thres_1pl <-
  prior("normal(0, 3)", class = "Intercept") +
  prior("normal(0, 3)", class = "sd", group = "id") 

formula <- brmsformula( y | thres(gr = item) ~ 1 + (1|id) )

fit2 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior_thres_1pl,
  cores = 8,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "../results/102-1pl-thres.rds"
)
toc()
```

```{r}
summary(fit2)
```

Extract diagnostic quantities:

```{r}
np <- nuts_params(fit2)
sum(subset(np, Parameter == "divergent__")$Value)
```

We compare the models using approximate LOO-CV:

```{r}
loo_compare(loo(fit1), loo(fit2))
```

We reject the assumption of only shifted thresholds.

# Fitting the ordered two-parameter IRT model

## Shifted thresholds

```{r}
tic("fit3")
prior <-
  # discrimination
  prior("normal(0, 1)", class = "Intercept", dpar = "disc") +             # alpha
  prior("normal(0, 1)", class = "sd", group = "item", dpar = "disc") +    # sigma_alpha
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 1)", class = "sd", coef = "Intercept", group = "item") 
  
formula <- brmsformula(
  y ~ 1 + (1|i|item) + (1|id),
  disc ~ 1 + (1|i|item)
)

fit3 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 8,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "../results/103-2pl.rds"
)
toc()
```
```{r}
prior_summary(fit3)
```

```{r}
summary(fit3)
```

Extract diagnostic quantities:

```{r}
np <- nuts_params(fit3)
sum(subset(np, Parameter == "divergent__")$Value)
```

## Item-specific thresholds

We add a discrimination parameter and maintain item-specific thresholds.

```{r}
tic("fit4")
prior_thres_2pl <-
  prior("normal(0, 3)", class = "Intercept") +
  prior("normal(0, 3)", class = "sd", group = "id") +
  prior("normal(0, 3)", class = "sd", group = "item", dpar = "disc")
  
formula <- brmsformula( y | thres(gr = item) ~ 1 + (1|id),
                        disc ~ (1|item))

fit4 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior_thres_2pl,
  cores = 8,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "../results/104-2pl-thres.rds"
)
toc()
```

```{r}
prior_summary(fit4)
```

```{r}
summary(fit4)
```

Extract diagnostic quantities:

```{r}
np <- nuts_params(fit4)
sum(subset(np, Parameter == "divergent__")$Value)
```

## Model comparision 

```{r}
loo_compare(loo(fit1), loo(fit2), loo(fit3), loo(fit4))
```

# Interpretation

## Easiness and discrimination from two-parameter IRT model with shifted thresholds (model 3)

```{r}
# extract item and easiness parameters
ranef1 <- ranef(fit3)
# discriminations
alpha <- ranef1$item[, , "disc_Intercept"] %>%
  exp() %>% 
    as_tibble() %>%
    rownames_to_column()
# item easinesses (deviations from thresholds)
beta <- ranef1$item[, , "Intercept"] %>%
    as_tibble() %>%
    rownames_to_column()
# put easinesses and discriminations together
bind_rows(beta, alpha, .id = "nlpar") %>%
    rename(item = "rowname") %>%
    mutate(item = as.numeric(item)) %>%
    mutate(nlpar = factor(nlpar, labels = c("Easiness", "Discrimination"))) %>%
    ggplot(aes(item, Estimate, ymin = Q2.5, ymax = Q97.5)) +
    facet_wrap("nlpar", scales = "free_x") +
    geom_pointrange() +
    coord_flip() +
    labs(x = "Item Number") + 
  scale_x_discrete( limits=1:9 )
```

## Discrimination from two-parameter IRT model with item-specific thresholds (model 4)

```{r}
# extract item and easiness parameters
ranef1 <- ranef(fit4)
# discriminations
alpha <- ranef1$item[, , "disc_Intercept"] %>%
  exp() %>% 
    as_tibble() %>%
    rownames_to_column()
# # item easinesses (deviations from thresholds)
# beta <- ranef1$item[, , "Intercept"] %>%
#     as_tibble() %>%
#     rownames_to_column()
# put easinesses and discriminations together
bind_rows(alpha, .id = "nlpar") %>%
    rename(item = "rowname") %>%
    mutate(item = as.numeric(item)) %>%
    mutate(nlpar = factor(nlpar, labels = c("Discrimination"))) %>%
    ggplot(aes(item, Estimate, ymin = Q2.5, ymax = Q97.5)) +
    facet_wrap("nlpar", scales = "free_x") +
    geom_pointrange() +
    coord_flip() +
    labs(x = "Item Number") + 
  scale_x_discrete( limits=1:9 )
```
  

```{r}
knitr::knit_exit()
```
