---
title: |
  | IRT model with indicators for six NFC-LFA categories
author: "Christoph Wunder"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab
    highlight: pygments
    toc: yes
    toc_depth: 4
    number_sections: yes
    # toc_float:
    #   smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '2'
# bibliography: /Users/Christoph/Documents/Uni/Literatur/lit_DB/litdb_sp.bib
# biblio-style: apalike
# link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, warning=FALSE)
rm(list=ls())
library(tidyverse)
library(haven)
library(brms)
library(bayesplot)
library(ggplot2)
theme_set(theme_bw())
library(tictoc)
library(sampling)
library(shinystan)
```

This file estimates an ordinal IRT model with categorical predictors of agricultural sustainability. 

# Data

The data are prepared starting in line 155 of the dofile `200-nfc-v3.do`.


```{r}
d <- read_dta("../data/master.dta")
d <- rename(d, id = idn)
d$id <- as.integer(d$id)
d$item <- as.factor(d$item)
d$lfa <- as.factor(d$lfa)
d$y <- d$y + 1
d$y <- factor( d$y, levels = c("1", "2", "3", "4"), ordered = TRUE ) 
table(d$y)
```

Select subsample of data set:

```{r, eval=FALSE, include=FALSE}
# set.seed(647382)
# n <- round(length(unique(d$id))*0.05, 0)
# samplecluster <- cluster(d, clustername=c("id"), size = n, method="srswor")
# d <- getdata(d, samplecluster)
# glimpse(d, width = 60)
```

Generate dummies for energy and industrial:

```{r}
d <- d %>% mutate( nfc = case_when( 
  energy==0 & industrial==0 & lfa==0 ~ "none0",
  energy==0 & industrial==0 & lfa==1 ~ "none1",
  energy>0  & industrial==0 & lfa==0 ~ "enrg0",
  energy>0  & industrial==0 & lfa==1 ~ "enrg1",
  energy==0 & industrial>0  & lfa==0 ~ "ind0",
  energy==0 & industrial>0  & lfa==1 ~ "ind1") )
d$nfc <- relevel( factor(d$nfc), ref="none0")
```


# Descriptive statistics 

```{r}
nrow(d)
nrow(d)/9
```

```{r}
summary(d)
```

```{r}
ggplot( d, aes(x=y) ) +
  stat_count(width = 0.5) +
  xlab("") +
  facet_wrap(~item)
```

# Fitting the 2PL ordinal model for categories defined by NFC and LFA

```{r, warning=FALSE, message=FALSE}
tic("fit0")
prior <-
  # discrimination
  prior("normal(0, 1)", class = "Intercept", dpar = "disc") +             # alpha
  prior("normal(0, 1)", class = "sd", group = "item", dpar = "disc") +    # sigma_alpha
  # unit-specific effects 
  prior("constant(1)", class = "sd", group = "id") +
  # item-specific effects 
  prior("normal(0, 1)", class = "sd", coef = "Intercept", group = "item") +
  # coefficients 
  prior("normal(0, 1)", class = "b")

formula <- brmsformula(
  y ~ 1 + (1|i|item) + (1|id) + nfc,
  disc ~ 1 + (1|i|item)
)

fit0 <- brm( 
  formula = formula,
  data = d,
  family = cumulative(link = "logit"),
  prior = prior,
  cores = 8,
  iter = 1000,
  warmup = 500,
  refresh = 0,
  file = "../results/120-2pl-catX.rds"
)
toc()
```

```{r}
summary(fit0)
```


# Interpretation

## Easiness and discrimination parameters

```{r}
# extract item and easiness parameters
ranef1 <- ranef(fit0)
# discriminations
alpha <- ranef1$item[, , "disc_Intercept"] %>%
  exp() %>% 
    as_tibble() %>%
    rownames_to_column()
# item easinesses (deviations from thresholds)
beta <- ranef1$item[, , "Intercept"] %>%
    as_tibble() %>%
    rownames_to_column()
# put easinesses and discriminations together
bind_rows(beta, alpha, .id = "nlpar") %>%
    rename(item = "rowname") %>%
    mutate(item = as.numeric(item)) %>%
    mutate(nlpar = factor(nlpar, labels = c("Easiness", "Discrimination"))) %>%
    ggplot(aes(item, Estimate, ymin = Q2.5, ymax = Q97.5)) +
    facet_wrap("nlpar", scales = "free_x") +
    geom_pointrange() +
    coord_flip() +
    labs(x = "Item Number") + 
  scale_x_discrete( limits=1:9 )
```

## Predicted probabilities

```{r}
conditional_effects(fit0,  effects = c("nfc"), categorical = TRUE, prob = 0.8)
```

```{r}
knitr::knit_exit()
```
